---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
 [![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/checkr.svg?branch=master)](https://travis-ci.org/poissonconsulting/checkr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/checkr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/checkr)
[![Coverage Status](https://img.shields.io/codecov/c/github/poissonconsulting/checkr/master.svg)](https://codecov.io/github/poissonconsulting/checkr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/checkr)](https://cran.r-project.org/package=checkr)

# checkr

`checkr` is an R package to check the dimensions, classes, values, and names of scalars, vectors, lists and data frames.
The various functions provide informative errors (or warnings) that allow users to quickly identify and fix any problems.

The following code demonstrates its use
```{r, error = TRUE}
library(checkr)

# the starwars data frame in the dplyr package fails many of these checks
check_data(dplyr::starwars, values = list(
  height = c(66L, 264L),
  name = "",
  mass = c(20,1358, NA),
  hair_color = c("blond", "brown", "black", NA),
  gender = c("male", "female", "hermaphrodite", "none", NA)), 
    order = TRUE, nrow = c(81, 84), key = "hair_color", error = FALSE)
```

The two other main functions are `check_vector()` and `check_list()`.

```{r, error=TRUE}
y <- c(2,1,0,1,NA)
check_vector(y, values = 1:10, length = 2, unique = TRUE, sorted = TRUE, named = TRUE, error = FALSE)
```

### Values

The values argument can be used to check the values of a vector, element of a list or column of a data frame.

#### Class

To check the class simply pass an object of the desired class.
```{r, error = TRUE}
check_vector(y, values = numeric(0))
check_vector(y, values = integer(0))
```

#### Missing Values

To check that a vector does not include missing values pass a single non-missing value (of the correct class).
```{r, error = TRUE}
check_vector(y, 1)
```

To allow it to include missing values include a missing value.
```{r}
check_vector(y, c(1, NA))
```

And to check that it only includes missing values only pass a missing value (of the correct class)
```{r, error = TRUE}
check_vector(y, NA_real_)
```

#### Range

To check the range of a vector pass two non-missing values (as well as the missing value if required).
```{r, error = TRUE}
check_vector(y, c(0, 2, NA))
check_vector(y, c(-1, -10, NA))
```

#### Specific Values

To check the vector only includes specific values pass three or more non-missing values.
```{r, error = TRUE}
check_vector(y, c(0, 1, 2, NA))
check_vector(y, c(1, 1, 2, NA))
```

### Naming Objects

By default, the name of an object is determined from the function call.
```{r, error=TRUE}
check_vector(list(x = 1))
```

This simplifies things but results in less informative error messages when used in a pipe.
```{r, error = TRUE}
library(magrittr)
y %>% check_list()
```

The argument `x_name` can be used to override the name.
```{r, error = TRUE}
y %>% check_list(x_name = "y")
```

## Citation

```{r, comment="", echo=FALSE}
citation(package = 'checkr')
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/checkr/issues).

[Pull requests](https://github.com/poissonconsulting/checkr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

[datacheckr](https://github.com/poissonconsulting/datacheckr)

