---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/checkr.svg?branch=master)](https://travis-ci.org/poissonconsulting/checkr)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/checkr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/checkr)
[![Coverage Status](https://img.shields.io/codecov/c/github/poissonconsulting/checkr/master.svg)](https://codecov.io/github/poissonconsulting/checkr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/checkr)](https://cran.r-project.org/package=checkr)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/checkr)]( https://CRAN.R-project.org/package=checkr)
[![JOSS](http://joss.theoj.org/papers/10.21105/joss.00624/status.svg)](https://doi.org/10.21105/joss.00624)

# checkr

`checkr` is a light-weight R package of expressive, assertive, pipe-friendly functions to check the properties of common R objects.

In the case of failure the functions, which are designed to be used in scripts and packages,
issue informative error messages.

For an overview of the functions see the `checkr-naming` vignette and for a comparison with similar packages see the `assertive-programming` vignette.

## Demonstration

The following code demonstrates the `check_data()` function
```{r, error = TRUE}
library(checkr)

# the starwars data frame in the dplyr package fails many of these checks
check_data(dplyr::starwars, values = list(
  height = c(66L, 264L),
  name = "",
  mass = c(20,1358, NA),
  hair_color = c("blond", "brown", "black", NA),
  gender = c("male", "female", "hermaphrodite", "none", NA)), 
    order = TRUE, nrow = c(81, 84), key = "hair_color", error = FALSE)
```

## Syntax

`checkr` uses objects to check 
the values of other objects using an elegant and expressive syntax.

#### Class

To check the class simply pass an object of the desired class.
```{r, error = TRUE}
y <- c(2,1,0,1,NA)
check_vector(y, values = numeric(0))
check_vector(y, values = integer(0))
```

#### Missing Values

To check that a vector does not include missing values pass a single non-missing value (of the correct class).
```{r, error = TRUE}
check_vector(y, 1)
```

To allow it to include missing values include a missing value.
```{r}
check_vector(y, c(1, NA))
```

And to check that it only includes missing values only pass a missing value (of the correct class)
```{r, error = TRUE}
check_vector(y, NA_real_)
```

#### Range

To check the range of a vector pass two non-missing values (as well as the missing value if required).
```{r, error = TRUE}
check_vector(y, c(0, 2, NA))
check_vector(y, c(-1, -10, NA))
```

#### Specific Values

To check the vector only includes specific values pass three or more non-missing values or 
set `only  = TRUE`.
```{r, error = TRUE}
check_vector(y, c(0, 1, 2, NA))
check_vector(y, c(1, 1, 2, NA))
check_vector(y, c(1, 2, NA), only = TRUE)
```

## Naming Objects

By default, the name of an object is determined from the function call.
```{r, error=TRUE}
check_vector(list(x = 1))
```

This simplifies things but results in less informative error messages when used in a pipe.
```{r, error = TRUE}
library(magrittr)
y %>% check_list()
```

The argument `x_name` can be used to override the name.
```{r, error = TRUE}
y %>% check_list(x_name = "y")
```

## Scalars

The four wrapper functions 
`check_lgl()`, `check_int()`, `check_dbl()` and `check_str()` check whether an object 
is an attribute-less non-missing scalar logical (flag), integer, double (number) or character (string).
They are really useful for checking the types of arguments in functions

```{r, error = TRUE}
fun <- function(x) { check_lgl(x)}
fun(x = NA)
fun(x = TRUE)
fun(x = 1)
```

Additional scalar wrappers are `check_date()` and `check_dttm()` for scalar Date and POSIXct objects.
Alternatively you can roll your own using the more general `check_scalar()` function.

## Installation

To install the latest official release from [CRAN](https://CRAN.R-project.org/package=checkr)
```
install.packages("checkr")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/checkr)
```
install.packages("devtools")
devtools::install_github("poissonconsulting/err")
devtools::install_github("poissonconsulting/checkr")
```

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("checkr")
```

## Citation

```{r, comment="", echo=FALSE}
citation(package = "checkr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/checkr/issues).

[Pull requests](https://github.com/poissonconsulting/checkr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

[datacheckr](https://github.com/poissonconsulting/datacheckr)
