---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![stability-unstable](https://img.shields.io/badge/stability-unstable-yellow.svg)](https://github.com/joethorley/stability-badges#unstable)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/checkr.svg?branch=master)](https://travis-ci.org/poissonconsulting/checkr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/checkr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/checkr)
[![Coverage Status](https://img.shields.io/codecov/c/github/poissonconsulting/checkr/master.svg)](https://codecov.io/github/poissonconsulting/checkr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/checkr)](https://cran.r-project.org/package=checkr)

# checkr

`checkr` in an R package to informatively check the dimensions, classes, values, and names of vectors, lists and data frames.

## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("poissonconsulting/checkr")
```

## Demonstration

```{r warning=TRUE}
library(checkr)
x <- c(2,1,0,1,NA)

check_vector(x)

check_vector(x, values = 1:10, max_length = 2, unique = TRUE, sorted = TRUE, named = TRUE, error = FALSE)

check_vector("term", c("term1", "term2"), error = FALSE)

y <- -1
checkor(check_count(y), check_flag(y), error = FALSE)
```

## Inspiration

[datacheckr](https://github.com/poissonconsulting/datacheckr)

## Contribution

Please report any [issues](https://github.com/poissonconsulting/checkr/issues).

[Pull requests](https://github.com/poissonconsulting/checkr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
