---
title: "Packages for Assertive Programming"
author: "Joe Thorley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Packages for Assertive Programming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Assertive programming follows the principles of fail fast and fail visibly.
It is implemented by issuing an informative error message if the function arguments fail to satisfy specific criteria.
This is particularly important in R because it is a dynamically typed language.

There are a host of R packages to facilitate assertive programming.
Here I am interested in those packages that are

1) on CRAN;
2) lightweight (few or no dependencies);
3) can be used to test the types and values of common R objects (not just data frames);
4) issue error messages.

The most useful packages that I am aware of that satisfy these criteria are

- [base](https://github.com/hadley/assertthat)
- [assertthat](https://github.com/hadley/assertthat)
- [checkmate](https://github.com/mllg/checkmate)
- [checkr](https://github.com/poissonconsulting/checkr)

This vignette consists of an overview of each package followed by a comparison.

In short the `checkr` package is recommended over the alternatives if you are looking for a set of expressive, light-weight, pipe-friendly assertive functions with customizable object names.

## Package Overview

### base

The `base` package offers the `stopifnot()` function.

```{r, error = TRUE}
height <- dplyr::starwars$height
stopifnot(is.character(height))
```
The error message is not that helpful if the user doesn't know what `is.character()` tests for.

In a function it behaves as follows:
```{r, error = TRUE}
my_fun <- function(x) stopifnot(is.character(x))
my_fun(height)
```

### assertthat

The [`assertthat`](https://github.com/hadley/assertthat) package offers `assert_that()`.
```{r, error=TRUE}
library(assertthat)
assert_that(is.character(height))
```
The error message is a big improvement on `stopifnot()`'s.

In a function it behaves as follows:
```{r, error = TRUE}
my_fun <- function(x) assert_that(is.character(x))
my_fun(dplyr::starwars$height)
```

Perhaps the neatest feature is the ability to set your own error message for an assertion function using `on_failure()`.
```{r, error = TRUE}
is.character2 <- function(x) is.character(x)
on_failure(is.character2) <- function(call, env) {
  paste0(deparse(call$x), " must be a character - sort it out!")
}
assert_that(is.character2(height))
```

```{r, echo = FALSE}
detach("package:assertthat")
```

### checkmate

The [`checkmate`](https://github.com/mllg/checkmate) package provides a plethora of functions to check the type and related properties of common objects.

```{r, error = TRUE}
library(checkmate)
assert_character(height)
```

In a function it behaves as follows
```{r, error = TRUE}
my_fun <- function(x) assert_character(x)
my_fun(height)
```

In addition, the `assert()` function allows multiple alternative `check**` functions (those which return a string of the error message on failure and a `TRUE` on success) to be combined into one assertion.
```{r, error = TRUE}
assert(checkCharacter(height), checkFactor(height))
```

```{r, echo = FALSE}
detach("package:checkmate")
```

### checkr

The [`checkr`](https://github.com/poissonconsulting/checkr) package provides a set of expressive functions to test the values of objects.

```{r, error = TRUE}
library(checkr)
check_vector(height, "")
```

In a function it behaves as follows:
```{r, error = TRUE}
my_fun <- function(x) check_vector(x, "")
my_fun(height)
```

And multiple alternative checks can be combined using the `checkor()` function.
```{r, error = TRUE}
checkor(check_vector(height, ""), check_vector(height, 1))
```

## Package Comparison

### Return Values

If all checks are passed `stopifnot()` returns an invisible `NULL`,
while `assert_that()` returns `TRUE`.
In contrast, `checkmate::assert_**` and `checkr::check_**`
return a copy of the original object which allows them to be used in pipes.

### Custom Error Messages

Although there is no way to customize the error message produced by `stopifnot()`,
`assert_that()` allows the programmer to specify the message for a individual test using the `msg` argument.
In the case of `checkmate` and `checkr` the programmer can specify the variable name using the `.var.name` and `x_name` argument, respectively. This can be very helpful in functions.
```{r, error = TRUE}
my_fun <- function(x) check_vector(x, "", x_name = substitute(x))
my_fun(height)
```

### Speed

Much of the `checkmate` package is written in `C` to minimize execution time.

### Dependencies

Whereas the `checkmate` package depends on `backports` and `utils`,
and `assertthat` depends on `tools`, and `checkr` depends on `err` only `base` is dependency-free.
