---
title: "checkr: An R package to check object values"
authors:
- name: Joe Thorley
  orcid: 0000-0002-7683-4592
  affiliation: 1
output:
  html_document:
    keep_md: yes
tags:
- R
affiliations:
  index: 1
  name: Poisson Consulting, Nelson, British Columbia
date: "`r Sys.Date()`"
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, messages = FALSE, fig.width = 8, fig.height = 12)
```

## Introduction

`checkr` is a lightweight R package to check the dimensions, classes, values, and names of scalars, vectors, lists and data frames.
The various functions provide informative errors (or warnings) that allow users to quickly identify and fix any problems.

## Demonstration

The following code demonstrates its use
```{r, warning = TRUE}
library(checkr)

# the starwars data frame in the dplyr package fails many of these checks
check_data(dplyr::starwars, values = list(
  height = c(66L, 264L),
  name = "",
  mass = c(20,1358, NA),
  hair_color = c("blond", "brown", "black", NA),
  gender = c("male", "female", "hermaphrodite", "none", NA)), 
    order = TRUE, nrow = c(81, 84), key = "hair_color", error = FALSE)
```

The two other main functions are `check_vector()` and `check_list()`.

```{r, warning = TRUE}
y <- c(2,1,0,1,NA)
check_vector(y, values = 1:10, length = 2, unique = TRUE, sorted = TRUE, named = TRUE, error = FALSE)
```
